---
title: "Author Collaboration Analytics"
output: 
  flexdashboard::flex_dashboard:
    source_code: "https://github.com/Ifeanyi55/authorPub-analytics/tree/main"
    favicon: bar.png
    theme:
      bg: "white"
      fg: "black" 
      primary: "green"
      base_font:
        google: Prompt
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: fill
    
---

```{r setup, include=FALSE}
library(flexdashboard)
# Install thematic and un-comment for themed static plots (i.e., ggplot2)
# thematic::thematic_rmd()
```

At A Glance {data-icon="fa-bar-chart"}
=========================================================
Column {data-width=650 .tabset} 
-----------------------------------------------------------------------
### **Publications** 

```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

# publication with the most collaborators
pubs_with_most_collabs <- data |> 
  group_by(Publication) |> 
  summarize(num_of_collabs = n()) |> 
  arrange(desc(num_of_collabs))

# top 20 pie
pie_hc <- pubs_with_most_collabs[1:20,] |> 
  hchart(
    type = "pie",
    hcaes(x = Publication,y = num_of_collabs),
    name = "Number of Authors",
    shadow = T) |> 
  hc_title(text = "<b>Top 20 Publications With The Most Collaborating Authors</b>")

pie_hc

```

### **Collaboration**

```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

# gauge plot of the highest number of authors
gauge <- highchart() |> 
  hc_pane(startAngle = -90,
          endAngle = 90,
          background = list(
            shape = "arc",
            innerRadius = "60%",
            outerRadius = "100%",
            borderWidth = 5
            )) |> 
  hc_yAxis(
    labels = list(style = list(fontSize = "15px",color = "black"))
  ) |> 
  hc_colors(colors = "orange") |> 
  hc_add_series(data = 54, 
                type = "solidgauge",
                name = "Number of Authors",
                dataLabels = list(
                  useHTML = T,
                  borderWidth = 0,
                  borderColor = "black",
                  style = list(fontSize = "20px",color = "green")
                )) |> 
  hc_tooltip(enabled = T) |> 
  hc_title(text = "<b>Highest Number of Authors Collaborating on One Publication</b>")

gauge

```

### **Authors**
```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

# top 20 authors involved in most collaboration
author_most_appearing <- data |> 
  group_by(Authors) |> 
  summarize(Frequency = n()) |> 
  arrange(desc(Frequency))

hc_bar <- author_most_appearing[1:20,] |> 
  hchart(
    type = "bar",
    hcaes(x = Authors,
          y = Frequency),
    shadow = T,
    name =  "Number of Collaborations",
    color = "#2ECC71"
  ) |> 
  hc_title(text = "<b>Top 20 Authors Involved in Most Collaborations</b>")

hc_bar

```

### **Position**

```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

most_author_position <- data |> 
  group_by(Position) |> 
  summarize(Count = n()) |> 
  arrange(desc(Count))

most_pie <- most_author_position |> 
  hchart(
    hcaes(x = Position, y = Count),
    type = "pie",
    name = "Position Count",
    shadow = T) |>
  hc_colors(c("#2874A6","#3498DB","#5DADE2")) |> 
  hc_title(text = "<b>Authors Positions Count</b>")

most_pie

```

Column {data-width=350}
-----------------------------------------------------------------------

### **Affiliations**

```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

# affiliations with most authors
aff_with_most_authors <- data |> 
  group_by(Affiliation) |> 
  summarize(num_of_auhtors = n()) |> 
  arrange(desc(num_of_auhtors))

# top 20 bar
hc <- aff_with_most_authors[1:20,] |> 
  hchart(type = "bar",
         hcaes(x = Affiliation, 
               y = num_of_auhtors),
         name = "Number of Authors",
         color = "#0174c3",
         shadow = T) |> 
  hc_title(text = "<b>Top 20 Affiliations With The Most Authors</b>")
  
hc

```

### **Institutions**

```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

# institutions with most authors
inst_with_most_authors <- data |> 
  group_by(Institution) |> 
  summarize(num_of_authors = n()) |> 
  arrange(desc(num_of_authors))


pie_hc2 <-  inst_with_most_authors[1:20,] |>
  hchart(
    hcaes(x = Institution, y = num_of_authors),
    type = "pie",
    name = "Number of Authors",
    shadow = T) |>     
  hc_title(text = "<b>Top 20 Institution Categories With The Most Authors</b>")

pie_hc2 

```

Narrative {.storyboard data-icon="fa-book"}
=========================================================


### The **Top 20 publications** with the most collaborating authors. {data-commentary-width=300}


```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

# publication with the most collaborators
pubs_with_most_collabs <- data |> 
  group_by(Publication) |> 
  summarize(num_of_collabs = n()) |> 
  arrange(desc(num_of_collabs))

# top 20 pie
pie_hc <- pubs_with_most_collabs[1:20,] |> 
  hchart(
    type = "pie",
    hcaes(x = Publication,y = num_of_collabs),
    name = "Number of Collaborating Authors",
    shadow = T) |> 
  hc_title(text = "<b>Top 20 Publications With The Most Collaborating Authors</b>") 
pie_hc

```

***
The data used in this project was scraped from [OpenAlex.org](https://openalex.org/) using keyword search. The keyword used to search for publications and their authors between the 1st of January 2020 and the 31st of August 2023 was ***simulation***.

The pie chart in the frame shows the top 20 publications in the data that has the most author collaborations. 

These are authors who worked together on the research project. 

As you hover over the pie chart, you will see the titles of their publications and the number of collaborating authors in the tooltip. 

Data Source: [OpenAlex.org](https://openalex.org/)

### The **highest number of authors** collaborating on a publication {data-commentary-width=300}

```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

# gauge plot of the highest number of authors
gauge <- highchart() |> 
  hc_pane(startAngle = -90,
          endAngle = 90,
          background = list(
            shape = "arc",
            innerRadius = "60%",
            outerRadius = "100%",
            borderWidth = 5
            )) |> 
  hc_yAxis(
    labels = list(style = list(fontSize = "15px",color = "black"))
  ) |> 
  hc_colors(colors = "orange") |> 
  hc_add_series(data = 54, 
                type = "solidgauge",
                name = "Number of Collaborating Authors",
                dataLabels = list(
                  useHTML = T,
                  borderWidth = 0,
                  borderColor = "black",
                  style = list(fontSize = "20px",color = "green")
                )) |> 
  hc_tooltip(enabled = T) |> 
  hc_title(text = "<b>Highest Number of Authors Collaborating on One Publication</b>")

gauge

```

***
The gauge plot shows the highest number of authors who collaborated on one publication.

The title of the publication that holds this distinction of having the most collaborating authors is ***Social determinants of mortality from COVID-19: A simulation study using NHANES***

Data Source: [OpenAlex.org](https://openalex.org/)

### The **Top 20 authors** involved in most collaborations {data-commentary-width=300}

```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

# top 20 authors involved in most collaboration
author_most_appearing <- data |> 
  group_by(Authors) |> 
  summarize(Frequency = n()) |> 
  arrange(desc(Frequency))

hc_bar <- author_most_appearing[1:20,] |> 
  hchart(
    type = "bar",
    hcaes(x = Authors,
          y = Frequency),
    shadow = T,
    name =  "Number of Collaborations",
    color = "#2ECC71"
  ) |> 
  hc_title(text = "<b>Top 20 Authors Involved in Most Collaborations</b>")

hc_bar

```

***
The bar plot shows the top 20 individual authors who have been involved in the most collaborations.

This implies that they have co-authored more than one publication in the data.

The author who holds the distinction for the most collaborations is ***Chunwei Zhang***.

Hover over the plot to see the collaborations count of the other authors.

Data Source: [OpenAlex.org](https://openalex.org/)

### The authors' **position count** as analyzed in the data {data-commentary-width=350}

```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

most_author_position <- data |> 
  group_by(Position) |> 
  summarize(Count = n()) |> 
  arrange(desc(Count))

most_pie <- most_author_position |> 
  hchart(
    hcaes(x = Position, y = Count),
    type = "pie",
    name = "Position Count",
    shadow = T) |>
  hc_colors(c("#2874A6","#3498DB","#5DADE2")) |> 
  hc_title(text = "<b>Authors Positions Count</b>")

most_pie

```

***
The pie chart shows the count of the authors' position in the collaborative authoring of the publications in the data.

The first author is the primary contributor who conducts research, analyzes data, and plays a central role in writing the paper, typically listed first and receiving significant credit.

Middle authors make substantial contributions, such as conducting experiments or providing expertise, with their order reflecting their relative contributions to the research.

The last author, often the senior author or mentor, provides guidance, secures resources, and oversees the project, listed at the end of the authorship order and associated with leadership.

Data Source: [OpenAlex.org](https://openalex.org/)

### The **Top 20 affiliations** with the most authors {data-commentary-width=300}

```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

# affiliations with most authors
aff_with_most_authors <- data |> 
  group_by(Affiliation) |> 
  summarize(num_of_auhtors = n()) |> 
  arrange(desc(num_of_auhtors))

# top 20 bar
hc <- aff_with_most_authors[1:20,] |> 
  hchart(type = "bar",
         hcaes(x = Affiliation, 
               y = num_of_auhtors),
         name = "Number of Authors",
         color = "#0174c3",
         shadow = T) |> 
  hc_title(text = "<b>Top 20 Affiliations With The Most Authors</b>")
  
hc

```

***
The bar plot shows the top 20 research affiliations having the most authors who collaborated on the publications in the data.

The leading affiliate organization is ***D-Wave Systems (Canada)***.

Data Source: [OpenAlex.org](https://openalex.org/)

### The **Top 20 institution categories** having the most authors {data-commentary-width=300}

```{r}
library(tidyverse)
library(highcharter)

# load authorPublications.csv
data <- read.csv("authorsPublications.csv")

# institutions with most authors
inst_with_most_authors <- data |> 
  group_by(Institution) |> 
  summarize(num_of_authors = n()) |> 
  arrange(desc(num_of_authors))


pie_hc2 <-  inst_with_most_authors[1:20,] |>
  hchart(
    hcaes(x = Institution, y = num_of_authors),
    type = "pie",
    name = "Number of Authors",
    shadow = T) |>     
  hc_title(text = "<b>Top 20 Institution Categories With The Most Authors</b>")

pie_hc2 


```

***
The pie chart shows institution categories having the most collaborating authors.

The top institution category is ***education***.

Data Source: [OpenAlex.org](https://openalex.org/)
